\textbf{On the Common Logical Structure of Classical and Quantum Mechanics}

### 2022-03-06 Articles/2020-QuantumLogic/Source/quantumlogic.tex
QuantumLogic: response to referee comments from Andrea

\author{Andrea Oldofredi\thanks{Section de Philosophie, Universit\'e de Lausanne, Switzerland. E-mail: Andrea.Oldofredi@unil.ch} \and Gabriele Carcassi\thanks{Physics Department, University of Michigan, Ann Arbor, MI 48109. E-mail: carcassi@umich.edu} \and Christine A. Aidala\thanks{Physics Department, University of Michigan, Ann Arbor, MI 48109. E-mail: caidala@umich.edu}}

@@

\begin{abstract}
It is generally believed that quantum mechanics entails a rejection of classical logic. The main arguments to support this claim come from the non-commutativity of quantum observables, which allegedly would generate a non-distributive lattice of propositions, and from quantum superpositions, which would entail new rules for quantum disjunctions. In this essay we show that such arguments contain serious flaws.  In particular, it is possible to keep the classical distributivity law in the context of quantum theory once the actual meaning of quantum propositions is unambiguously defined and properly taken into account. We moreover argue that the existence of quantum superpositions does not lead to the definition of a new rule for logical disjunctions. Furthermore, contrary to the received view in quantum logic, we explain that quantum mechanics can generate a distributive lattice of propositions. Finally, the remarkable formal similarities between the logical structures of classical and quantum mechanics are highlighted and discussed. 
\vspace{4mm}

@@

As soon as Quantum Mechanics (QM) achieved a definite and coherent mathematical formulation, several physicists and philosophers claimed that its formal structure does not conform to the laws of classical propositional calculus.\footnote{For an introduction and a clear discussion of quantum logic the reader may refer to cf.\ \cite{Giuntini:2002}, \cite{DallaChiara:2004}, \cite{deRonde:2016}, \cite{Wilce:2021}.} This belief stood the test of time, and it is still widespread to the present day. For example, \cite{Giuntini:2018} claim that ``[t]he \emph{logic} of classical physical objects is naturally based on a two-valued semantics. Such a dichotomic situation breaks down in quantum theory'' (p.\ 4), implying that the latter entails a revision of classical propositional calculus. Such a view is shared by \cite{Wilce:2021}, who says: ``[m]athematically, quantum mechanics can be regarded as a non-classical probability calculus resting upon a non-classical propositional logic''. 
\noindent Similarly, presenting the standard formulation of quantum theory and its peculiar features, Stefanovich states that
\begin{quote}
the true logical relationships between results of measurements are different from the classical laws of Aristotle and Boole. The usual classical logic needs to be replaced by the so-called quantum logic (\cite{Stefanovich:2019}, p.\ 2)
\end{quote}
\noindent Even more explicitly, introducing the quantum logical formalism Holik writes that
\begin{quote}
[t]he connectives ``$\vee$'', ``$\wedge$'' and ``$\neg$'' are the quantum logical analogues of the Boolean connectives. It is important to remark that, with these connectives, the lattice of closed subspaces does not satisfies [\emph{sic}] the distributivity law [...] and then, it is non-Boolean. This implies that the new connectives cannot be interpreted in a classical way (\cite{Holik:2021}, p.\ 5).
\end{quote}

The first modifications of Classical Logic (CL) were advanced in the thirties in order to faithfully represent the physical content of quantum theory. For instance, in 1931 the Polish philosopher Zygmunt Zawirski proposed to apply \L ukasiewicz's three-valued logic to QM, starting from considerations about the wave-particle duality and Heisenberg's uncertainty principle. In virtue of the latter, moreover, the logical empiricists Moritz Schlick and Philipp Frank claimed that the conjunction of two or more statements may have no meaning in quantum theory, concluding that the classical conjunction is not always valid in the context of quantum physics, as recalled in \cite{Carnap:1966}. In their view, a typical example of invalid conjunction is one attributing a definite value for both position and momentum to a single quantum particle at a given time $t$. Furthermore, in 1933 Fritz Zwicky suggested that QM rejects the law of excluded-middle.\footnote{For historical details on quantum logic see \cite{Jammer:1974}.} However, it is generally recognized that the founding text of Quantum Logic (QL) is Birkhoff and von Neumann's essay ``\emph{The Logic of Quantum Mechanics}'' (\cite{vonNeumann:1936}), where the authors provided the standard account of the propositional calculus obeyed by quantum propositions.\footnote{It is worth noting that von Neumann in his treatise \emph{Mathematische Grundlagen der Quantenmechanik} published in 1932 anticipated that from the algebraic structure of quantum theory it would have been possible to formulate a new propositional calculus. These ideas would be fully expressed in his successive collaboration with Garrett Birkhoff.}

A few years after the publication of this paper, the research in quantum logic saw a notable development in both the scientific and philosophical communities, opening new research programs searching for the correct logical structures to understand and describe the physics of QM (cf.\ \cite{Reichenbach:1944}, \cite{Mackey:1957}, \cite{Finkelstein:1963}, \cite{Kochen:1965}, \cite{Jauch:1969}, \cite{Giuntini:2002}, \cite{DallaChiara:2004}, \cite{Pitowsky:2006}, \cite{Giuntini:2018}, \cite{Svozil:2020}, \cite{Holik:2021}), and intense metaphysical debates regarding the nature of logic itself and its empirical status (cf.\ notably \cite{Quine:1951}, \cite{Putnam:1968}, \cite{Dummett:1976}, \cite{Hallett:1982}, \cite{Weingartner:2004}, \cite{Bacciagaluppi:2009}). In addition, it is worth noting that research in quantum logic is still very prolific to the present day with many interesting applications in the fields of quantum information, computation and cryptography. Therefore, although QL is no longer viewed as a solution to the foundational problems affecting quantum theory\footnote{This conclusion is now accepted among experts; for details see e.g.\ \cite{Bacciagaluppi:2009} and \cite{Giuntini:2002}. Referring to this, the latter authors stated explicitly that ``quantum logics are not to be regarded as a kind of ``clue'', capable of solving the main physical and epistemological difficulties of QT [quantum theory]. This was perhaps an illusion of some pioneering workers in quantum logic'' (\cite{Giuntini:2002}, p.\ 225).}, it remains an important discipline of contemporary physics. In this paper we will neither take a stance in the dispute concerning the ontological status of logic and its laws, nor we will consider recent developments of QL given that they are not strictly relevant to the point we are going to make. Indeed, the present essay is focused on the logical structures common to both quantum and classical mechanics, following the spirit of the former discussion.\footnote{In the present essay we will be concerned only with propositional calculus in the context of non-relativistic QM. First-order and higher-order logics will not be discussed here, nor the logic of relativistic quantum theories.}

Referring to this, one of us recently faced the question of whether quantum physics implies a revision of classical logic (cf.\ \cite{Oldofredi:2020}). Alternatively stated, it was asked if classical logic must be inevitably abandoned in the quantum realm, or if there are quantum theories compatible with classical propositional calculus. Analyzing Bohmian mechanics, a well-known interpretation of QM, it has been shown that quantum physics does not necessarily involve a departure from classical logic. More precisely, it has been argued that in this interpretation of QM not only the logical connectives retain their classical meanings---in particular the negation and the disjunction do follow classical rules---but also the distributivity law holds. This fact is a consequence of the particle ontology of the theory and its explanation of quantum measurements.

@@

% In the following section we will see that Hilbert spaces already posses a mathematical structure that follows classical logic, namely the Borel $\sigma$-algebra, and therefore quantum mechanics already posses a propositional lattice that follows classical logic. Moreover, we show that classical probabilistic theories also feature a non-distributive lattice that is the analogue of quantum logic, which means non-distributivity is neither required nor unique to quantum mechanics.

@@

In this section we will review the examples given in the previous one and explain that the failure of distributivity stems from the ambiguity of the quantum logical propositions and, once that is clarified, standard classical rules can be applied also in the quantum context. More precisely, we are going to argue that (i) it is not necessarily the case that a quantum disjunction must be true when neither of its members is, and (ii) that \eqref{DL} can be retained in QM, since the arguments seen above---usually meant to show its failure---are flawed.

@@

On the other hand, we can calculate the truth value of $q_i \vee r_i$ directly and show that it is false. The difficulty here is that before the measurement of $S_y$ only $x$-spin is defined, so how are we going to compute statements about the value of $y$-spin? The solution is simple: while the value of an observable is not always defined, its expectation value always is. The idea is then to translate $q_i$ and $r_i$ into propositions about expectation values, which we already know how to handle, and apply the standard rules. Consider the following statements:
\begin{description}
	\item $q'_i$: ``before the measurement, the expectation value for $y$-spin is $\frac{1}{2} \hbar$'',
	\item $r'_i$: ``before the measurement, the expectation value for $y$-spin is $-\frac{1}{2} \hbar$''.
\end{description}
\noindent Because the spin of the electron along any axis is bounded by $\pm \frac{1}{2} \hbar$, the only way that $q'_i$ is true is if $q_i$ is true. Therefore $q'_i$ and $r'_i$ are respectively equivalent to $q_i$ and $r_i$. Consequently $q_i \vee r_i = q'_i \vee r'_i$. Note that $q'_i \vee r'_i$  is the proposition ``before the measurement, the expectation value for $y$-spin is $\pm \frac{1}{2} \hbar$''. Now, before the measurement of $S_y$ we know that the electron $x$-spin is up. While the $y$-spin value is not fully defined---the state is not an eigenstate of $S_y$---the expectation value is fully defined and it is zero. Zero is different from $\pm \frac{1}{2} \hbar$ and therefore $q_i \vee r_i = \bot$. In fact, since any (non-trivial) superposition of $y$-spin up and down will result in an expectation value strictly between $-\frac{1}{2}\hbar$ and $\frac{1}{2}\hbar$, $q_i \vee r_i$ will be true if and only if either $q_i$ or $r_i$ is true. Therefore it is not the case that the quantum disjunction can be true even if none of the elements is true.

@@

Let us now turn our attention to distributivity and try to rewrite \eqref{contradiction} by using the refined propositions that explicitly include the temporal dependence. The proposition $p$ always referred to preparation, therefore it maps to $p_i$ in all cases. The disjunction was argued to be true on the basis of what happened \emph{after} the measurement of $S_y$, therefore on the left side of the relationship we should use $q_o$ and $r_o$. On the right side, instead, it was argued that $p \wedge q$ was false because of the incompatibility between the spin operators among different axis. This means we are considering statements at equal time with $p_i$, and therefore we should use $q_i$ and $r_i$. We have 
\begin{align}
	\label{solved_contradiction}
	p_i \wedge (q_o \vee r_o) = \top \longleftrightarrow (p_i \wedge q_i) \vee (p_i\wedge r_i) = \bot.
\end{align}
\noindent But this is hardly a failure of distributivity, since the statements on the left side are not the same as the statements on the right side. Let us fix that.

@@

\begin{figure}

\begin{center}
\includegraphics[scale=.8]{Ball.eps}
\end{center}
\caption{\small{Graphic representation of the discussed example.  Initially a ball is in position $p$ over a hatch.  After the hatch opens, the ball drops and falls to either position $q$ or $r$ with equal chance.}}
\end{figure}

@@

Since $\langle \psi , A \psi \rangle$ expresses the expected value for operator $A$, the proposition shows that the Borel algebra can express statements about the expectation value of observables,\footnote{Technically, position and momentum are not bounded operators. However, one can argue that measured position and momentum are indeed bounded as the acceptance of a detector is always limited. Note that all operators for which the spectral theorem applies can be written as the limit of a sequence of bounded operators.} such as ``the expected position $\langle x \rangle$ of the particle is between $x_0$ and $x_1$.'' This would correspond to the Borel set $\{ \psi \in \mathcal{H} \, | \, \langle \psi , A \psi \rangle \in (x_0, x_1) \}$.\footnote{In most cases, one assumes states to be normalized, which is what we have done here for simplicity. Renormalization would lead to the set $\{ \psi \in \mathcal{H} \, | \, \langle \psi , A \psi \rangle / \langle \psi , \psi \rangle \in (x_0, x_1) \}$, which is still a Borel set since renormalization is a continuous function over its domain $\mathcal{H} \setminus \{ 0 \}$.} Consider, for example, ``the expectation of $z$-spin is zero''. This would correspond to the set $V \subseteq \mathcal{H}$ that includes all states, and only the states, for which the expection of $S_z$ is zero. By Proposition 2, $V$ is a Borel set. In fact, the set $U = {0}$ is a Borel set of $\mathbb{R}$ and therefore $V = F_{S_z}^{-1}(U)$ is a Borel set of $\mathcal{H}$. However, $V$ is not a closed subspace. The set $V$ is not the full set $\mathcal{H}$ since, for example, the state with $z$-spin up is not in the set. However, $x$-spin up and $x$-spin down will be in the set $V$, and their span is the full space $\mathcal{H}$. Therefore $V$ is not a closed subspace. Our proposition, despite being physically interesting, does not correspond to a closed subspace but does correspond to a Borel set.

>>>>

\author{The Authors}

@@

\begin{abstract}
\textcolor{blue}{At the onset of quantum mechanics, it was argued that the new theory would entail a rejection of classical logic. The main arguments to support this claim come from the non-commutativity of quantum observables, which allegedly would generate a non-distributive lattice of propositions, and from quantum superpositions, which would entail new rules for quantum disjunctions. While the quantum logic program is not as popular as it once was, a crucial question remains unsettled: what is the relationship between the logical structures of classical and quantum mechanics? In this essay we answer this question by showing that the original arguments promoting quantum logic contain serious flaws, and that quantum theory does satisfy the classical distributivity law once the full meaning of quantum propositions is properly taken into account. Moreover, we show that quantum mechanics can generate a distributive lattice of propositions, which, unlike the one of quantum logic, includes statements about expectation values which are of undoubtable physical interest. Lastly, we show that the lattice of statistical propositions in classical mechanics follows the same structure, yielding an analogue non-commutative sublattice of classical propositions. This fact entails that the purported difference between classical and quantum logic stems from a misconstructed parallel between the two theories.}
\vspace{4mm}

@@

As soon as Quantum Mechanics (QM) achieved a definite and coherent mathematical formulation, several physicists and philosophers claimed that its formal structure does not conform to the laws of classical propositional calculus.\footnote{For an introduction and a clear discussion of quantum logic the reader may refer to cf.\ \cite{Giuntini:2002}, \cite{DallaChiara:2004}, \cite{deRonde:2016}, \cite{Wilce:2021}.} The first modifications of Classical Logic (CL) were advanced in the thirties in order to faithfully represent the physical content of quantum theory. For instance, in 1931 the Polish philosopher Zygmunt Zawirski proposed to apply \L ukasiewicz's three-valued logic to QM, starting from considerations about the wave-particle duality and Heisenberg's uncertainty principle. In virtue of the latter, moreover, the logical empiricists Moritz Schlick and Philipp Frank claimed that the conjunction of two or more statements may have no meaning in quantum theory, concluding that the classical conjunction is not always valid in the context of quantum physics, as recalled in \cite{Carnap:1966}. In their view, a typical example of invalid conjunction is one attributing a definite value for both position and momentum to a single quantum particle at a given time $t$. Furthermore, in 1933 Fritz Zwicky suggested that QM rejects the law of excluded-middle.\footnote{For historical details on quantum logic see \cite{Jammer:1974}.} However, it is generally recognized that the founding text of Quantum Logic (QL) is Birkhoff and von Neumann's essay ``\emph{The Logic of Quantum Mechanics}'' (\cite{vonNeumann:1936}), where the authors provided the standard account of the propositional calculus obeyed by quantum propositions.\footnote{It is worth noting that von Neumann in his treatise \emph{Mathematische Grundlagen der Quantenmechanik} published in 1932 anticipated that from the algebraic structure of quantum theory it would have been possible to formulate a new propositional calculus. These ideas would be fully expressed in his successive collaboration with Garrett Birkhoff.}

A few years after the publication of this paper, the research in quantum logic saw a notable development in both the scientific and philosophical communities, opening new research programs searching for the correct logical structures to understand and describe the physics of QM (cf.\ \cite{Reichenbach:1944}, \cite{Mackey:1957}, \cite{Finkelstein:1963}, \cite{Kochen:1965}, \cite{Jauch:1969}, \cite{Giuntini:2002}, \cite{DallaChiara:2004}, \cite{Pitowsky:2006}, \cite{Giuntini:2018}, \cite{Svozil:2020}, \cite{Holik:2021}), and intense metaphysical debates regarding the nature of logic itself and its empirical status (cf.\ notably \cite{Quine:1951}, \cite{Putnam:1968}, \cite{Dummett:1976}, \cite{Hallett:1982}, \cite{Weingartner:2004}, \cite{Bacciagaluppi:2009}).

The belief according to which QM entails a new logic stood the test of time, and it is still \textcolor{blue}{alive} to the present day. For example, \cite{Giuntini:2018} claim that ``[t]he \emph{logic} of classical physical objects is naturally based on a two-valued semantics. Such a dichotomic situation breaks down in quantum theory'' (p.\ 4), implying that the latter entails a revision of classical propositional calculus. Such a view is shared by \cite{Wilce:2021}, who says: ``[m]athematically, quantum mechanics can be regarded as a non-classical probability calculus resting upon a non-classical propositional logic''. \textcolor{blue}{Similarly, explaining the difference between the logic of classical mechanics and the logic obeyed by quantum propositions, \cite{Holik:2019} write that
\begin{quote}
[i]n classical and quantum mechanics, the physical properties of a system are endowed with a lattice structure. These structures are different in the classical
and quantum case, and they determine the logical structure of the physical system (\cite{Holik:2019}, p.\ 363-364).
\end{quote}
\noindent More precisely the authors say:
\begin{quote}
The distributive inequalities are the main difference between classical and quantum logic. In the classical lattice, all properties satisfy the distributive equalities, but in the quantum lattice, only distributive inequalities hold, in general (\emph{ibid.}).
\end{quote}
}

\textcolor{blue}{Even more explicitly, presenting the standard formulation of quantum theory and its peculiar features, Stefanovich states that
\begin{quote}
the true logical relationships between results of measurements are different from the classical laws of Aristotle and Boole. The usual classical logic needs to be replaced by the so-called quantum logic (\cite{Stefanovich:2019}, p.\ 2).
\end{quote}
}

 \textcolor{blue}{While the quantum logic programme is still being actively pursued by a sizable community, and despite the many interesting applications in the fields of quantum information, computation and cryptography that are being investigated to the present day, QL is no longer viewed as a solution to the foundational problems affecting quantum theory.\footnote{This conclusion is now accepted among experts; for details see e.g.\ \cite{Bacciagaluppi:2009} and \cite{Giuntini:2002}. Referring to this, the latter authors stated explicitly that ``quantum logics are not to be regarded as a kind of ``clue'', capable of solving the main physical and epistemological difficulties of QT [quantum theory]. This was perhaps an illusion of some pioneering workers in quantum logic'' (\cite{Giuntini:2002}, p.\ 225).} Nevertheless, although it may be reasonable for a physicist to simply discard the approach over practical considerations, such as the lack of a desired result, on philosophical grounds this move would beg important questions, as for instance: what is the relationship between the logical structures of classical and quantum mechanics? How are they different and how are they similar? What should we think about the core insights that led to the development of QL? Were they substantially correct, or should they be rejected? We think that providing a correct answer to these issues will lead (i) to a deeper understanding of the structural relations existing between classical logic and quantum mechanics---which unfortunately are not yet precisely understood---and (ii) to a clarification of important confusions concerning the motivations behind QL. %For these reasons, the present essay is focused on the logical structures common to both quantum and classical mechanics, highlighting the similarities between these two theoretical frameworks.
 \footnote{In the present essay we will be concerned only with propositional calculus in the context of non-relativistic QM. First-order and higher-order logics will not be discussed here, nor the logic of relativistic quantum theories.}}

Referring to this, one of us recently faced the question of whether quantum physics implies a revision of classical logic (cf.\ [Reference Deleted]). Alternatively stated, it was asked if classical logic must be inevitably abandoned in the quantum realm, or if there are quantum theories compatible with classical propositional calculus. Analyzing Bohmian mechanics, a well-known interpretation of QM, it has been shown that quantum physics does not necessarily involve a departure from classical logic. More precisely, it has been argued that in this interpretation of QM not only the logical connectives retain their classical meanings---in particular the negation and the disjunction do follow classical rules---but also the distributivity law holds. This fact is a consequence of the particle ontology of the theory and its explanation of quantum measurements.

@@

@@

In this section we will review the examples given in the previous one and explain that the failure of distributivity stems from the ambiguity of the quantum logical propositions and, once that is clarified, standard classical rules can be applied also in the quantum context. \textcolor{blue}{More precisely, we are going to argue that (i) it is not necessarily the case that a quantum disjunction can be true when neither of its members is\footnote{\textcolor{blue}{Clearly, a quantum disjunction is false when both disjuncts are false; however, the interesting case for our discussion is the one in which quantum propositions have undetermined truth values.}}---i.e.\ when both disjuncts have undetermined truth values as in the case discussed above in which we considered a system in a linear combination of $y$-spin states}---and (ii) that \eqref{DL} can be retained in QM, since the arguments seen above---usually meant to show its failure---are flawed.

@@

On the other hand, we can calculate the truth value of $q_i \vee r_i$ directly and show that it is false. The difficulty here is that before the measurement of $S_y$ only $x$-spin is defined, so how are we going to compute statements about the value of $y$-spin? The solution is simple: while the value of an observable is not always defined, its expectation value always is. The idea is then to translate $q_i$ and $r_i$ into propositions about expectation values, which we already know how to handle, and apply the standard rules. Consider the following statements:
\begin{description}
	\item $q'_i$: ``before the measurement, the expectation value for $y$-spin is $\frac{1}{2} \hbar$'',
	\item $r'_i$: ``before the measurement, the expectation value for $y$-spin is $-\frac{1}{2} \hbar$''.
\end{description}
\noindent \textcolor{blue}{If $\rho_i$ is the (potentially) mixed state before the measurement, the expectation value for $y$-spin is given by $\text{Tr}(\rho_i S_y) = \frac{1}{2} \hbar \langle \psi_y^+ | \rho_i | \psi_y^+ \rangle - \frac{1}{2} \hbar \langle \psi_y^- | \rho_i | \psi_y^- \rangle$. Therefore $q'_i$ is true if and only if $\text{Tr}(\rho_i S_y) = \frac{1}{2} \hbar$. Because $\rho_i$ is positive definite, this can only happen if $\langle \psi_y^+ | \rho_i | \psi_y^+ \rangle = 1$ and $\langle \psi_y^- | \rho_i | \psi_y^- \rangle = 0$, which means $\rho_i = | \psi_y^+ \rangle \langle \psi_y^+ |$. This entails that the expectation value for $y$-spin is $\frac{1}{2} \hbar$ if and only if the electron was prepared in a pure state of $y$-spin up. In other words, because the spin of the electron along any axis is bounded by $\pm \frac{1}{2} \hbar$, the only way that $q'_i$ is true is if $q_i$ is true. We have that $q'_i$ and $r'_i$ are respectively equivalent to $q_i$ and $r_i$.} Consequently $q_i \vee r_i = q'_i \vee r'_i$. Note that $q'_i \vee r'_i$ is the proposition ``before the measurement, the expectation value for $y$-spin is $\pm \frac{1}{2} \hbar$''. Now, before the measurement of $S_y$ we know that the electron $x$-spin is up. While the $y$-spin value is not fully defined---the state is not an eigenstate of $S_y$---the expectation value is fully defined and it is zero. Zero is different from $\pm \frac{1}{2} \hbar$ and therefore $q_i \vee r_i = \bot$. In fact, since any (non-trivial) superposition of $y$-spin up and down will result in an expectation value strictly between $-\frac{1}{2}\hbar$ and $\frac{1}{2}\hbar$, $q_i \vee r_i$ will be true if and only if either $q_i$ or $r_i$ is true. Therefore it is not the case that the quantum disjunction can be true even if none of the elements is true. \textcolor{blue}{The overall point is that the incompatibility of the observables does indeed introduce a problem because it is not clear how to evaluate propositions that mix different variables. The use of equivalent propositions over expectation values helps us solve the ambiguities.}

@@

\textcolor{blue}{Several authors have already argued in different ways that it is possible to retain distributivity in quantum mechanics. For instance, \cite{Park:1968} disputed the identification between Hermitian operators and quantum observables and provided a new interpretation of quantum measurement theory rejecting the notion of incompatible observables---and thereby incompatible measurements---which, as we have seen, constitutes the basis to show the failure of the distributivity law in the quantum realm. Remarkably, also Bohr argued that classical logic should be maintained in quantum theory as a consequence of the complementarity principle:
\begin{quote}
The aim of the idea of complementarity was to allow of keeping the usual logical forms while procuring the extension necessary for including the new situation relative to the problem of observation in atomic physics (Bohr quoted in \cite{Faye:2021}, p.\ 115).
\end{quote}
Another interesting quote reported in \cite{Faye:2021} clearly shows that the Danish physicists strongly opposed the idea of replacing classical propositional calculus with QL: 
\begin{quote}
The question has been raised whether recourse to multivalued logics is needed for a more appropriate representation of the situation. From the preceding argumentation it will appear, however, that all departures from common language and ordinary logic are entirely avoided by reserving the word `phenomenon' solely for reference to unambiguously communicable information, in the account of which the word `measurement' is used in its plain meaning of standardized comparison (Bohr quoted in \cite{Faye:2021}, p.\ 115).
\end{quote}
Although Bohr did not provide formal arguments against the introduction of quantum logic, his views express the idea according to which macroscopic measurement results must be described with ordinary language, therefore, they must be subjected to the rules of classical logic---in his view the intrinsic novelties of quantum mechanics have to be found in the contextual and complementary nature of quantum phenomena on the one hand, and the entanglement between quantum systems and macroscopic devices which in measurement situations form an indissoluble unity on the other. Hence, given that quantum propositions concern only measurement results, Bohr concluded that they should be governed by the laws of classical propositional calculus.
To this regard, \cite{Heelan:1970} provided a formalization of Bohr's arguments based on the complementarity principle and showed that the statements about a single quantum mechanical event, i.e.\ a measurement of a certain observable given a precise experimental context, do follow a classical propositional calculus.}\footnote{\textcolor{blue}{We thank one anonymous reviewer for having pointed out Heelan's work to us.}} \textcolor{blue}{More precisely, Heelan claims that QM introduces a distinction between events, corresponding to the actual performance of a certain measurement, and physical contexts, i.e.\ experimental settings determining the necessary and sufficient conditions for the realization of a particular measurement outcome. In his view, then, one has to introduce two different languages: an event-language useful to formally describe a particular observation of a ccertain quantum observable and a context-language, which is a meta-language necessary in order to properly speak about how a certain event can possibly occur.\footnote{\textcolor{blue}{Heelan exemplifies such a distinction as follows: ``If the event, for example, is a particle-location event, the event-language is position-language, and the physical context is a standardized instrumental set-up plus whatever other physical conditions are necessary and sufficient for the measurement of a given range of possible particle-position events'' (\cite{Heelan:1970}, p.\ 96).}} If we consider only propositions involving and relating incompatible contexts, Heelan says, then we will generate a lattice of propositions which is not distributive, however, if we take into account single event-languages---each of which is coupled with a given context-language---the logic of such individual quantum events can be classical. This captures Bohr's idea according to which the principle of complementarity arises from the context-dependent character of quantum mechanical events.}

\textcolor{blue}{Although we partially agree with Heelan's view in saying that propositions about individual quantum measurements can generate a distributive lattice, we want to generalize such a claim. In fact, we show---with simpler arguments and independently of Bohr's interpretation of quantum theory---that taking into account the temporal order of individual quantum measurements one can relate even statements about incompatible observations into a distributive lattice. In addition, we show that ignoring the temporal order of measurements may lead to violations of the distributivity law also in classical mechanics; thus, it is our aim to argue that once the correct interpretation of quantum propositions is taken into account the lattice generated by such statements is distributive.}

In order to derive our desired result, let us now turn our attention to distributivity and try to rewrite \eqref{contradiction} by using the refined propositions that explicitly include the temporal dependence. The proposition $p$ always referred to preparation, therefore it maps to $p_i$ in all cases. The disjunction was argued to be true on the basis of what happened \emph{after} the measurement of $S_y$, therefore on the left side of the relationship we should use $q_o$ and $r_o$. On the right side, instead, it was argued that $p \wedge q$ was false because of the incompatibility between the spin operators among different axis. This means we are considering statements at equal time with $p_i$, and therefore we should use $q_i$ and $r_i$. We have 
\begin{align}
	\label{solved_contradiction}
	p_i \wedge (q_o \vee r_o) = \top \longleftrightarrow (p_i \wedge q_i) \vee (p_i\wedge r_i) = \bot.
\end{align}
\noindent But this is hardly a failure of distributivity, since the statements on the left side are not the same as the statements on the right side. Let us fix that.

@@

\begin{center}
\includegraphics[scale=.8]{Ball.eps}
\begin{quote}
\small{Figure 2: Graphic representation of the discussed example.  Initially a ball is in position $p$ over a hatch.  After the hatch opens, the ball drops and falls to either position $q$ or $r$ with equal chance.}
\end{quote}
\end{center}

@@

Since $\langle \psi , A \psi \rangle$ expresses the expected value for operator $A$, the proposition shows that the Borel algebra can express statements about the expectation value of observables,\footnote{Technically, position and momentum are not bounded operators. However, one can argue that measured position and momentum are indeed bounded as the acceptance of a detector is always limited. Note that all operators for which the spectral theorem applies can be written as the limit of a sequence of bounded operators.} such as ``the expected position $\langle x \rangle$ of the particle is between $x_0$ and $x_1$.'' This would correspond to the Borel set $\{ \psi \in \mathcal{H} \, | \, \langle \psi , A \psi \rangle \in (x_0, x_1) \}$.\footnote{In most cases, one assumes states to be normalized, which is what we have done here for simplicity. Renormalization would lead to the set $\{ \psi \in \mathcal{H} \, | \, \langle \psi , A \psi \rangle / \langle \psi , \psi \rangle \in (x_0, x_1) \}$, which is still a Borel set since renormalization is a continuous function over its domain $\mathcal{H} \setminus \{ 0 \}$.} Consider, for example, ``the expectation of $z$-spin is zero''. This would correspond to the set $V \subseteq \mathcal{H}$ that includes all states, and only the states, for which the expection of $S_z$ is zero. By Proposition 2, $V$ is a Borel set. In fact, the set $U = {0}$ is a Borel set of $\mathbb{R}$ and therefore $V = F_{S_z}^{-1}(U)$ is a Borel set of $\mathcal{H}$. However, $V$ is not a closed subspace. The set $V$ is not the full set $\mathcal{H}$ since, for example, the state with $z$-spin up is not in the set. However, $x$-spin up and $x$-spin down will be in the set $V$, and their span is the full space $\mathcal{H}$. Therefore $V$ is not a closed subspace. Our proposition, despite being physically interesting, does not correspond to a closed subspace but does correspond to a Borel set.\footnote{\textcolor{blue}{Note that we are not claiming that all Borel sets can be constructed in this way or that all Borel sets correspond to physically interesting propositions. First of all, it is not true: $\{ \psi \}$ and $\{ e^{\imath \theta} \psi\}$ are different Borel sets, yet they are not physically distinguishable since a difference in absolute phase is not physically relevant. This can be fixed by considering the Borel sets of the projective space. Second, to really examine the physicality of all Borel sets we need a ``general theory of experimental logic'', of the type provided by \cite{Kelly:1996} or \cite{Carcassi:2021}, which would go well beyond the present discussion. With appropriate caveats (e.g. the Hilbert space must be separable), one can say that the Borel sets correspond to propositions that can be associated to an experimental test (what \cite{Carcassi:2021} call ``theoretical statements'') which may or may not terminate in any or all cases. We leave this discussion for another work.}}
